---
title: "Anoh_R_Practice 3"
output:
  word_document: default
  pdf_document: default
date: "2022-11-19"
---





##1- Conduct a one-sample t-tests for mean using an appropriate variable from the data set. Use the t.test( ) in R. Provide the null and the alternative hypothesis for each test. Provide the test results and several sentences interpreting the results.


```{r}
library(dplyr)
library(tidyverse)
library(plotrix) 
library(readr)
library(ggpubr)
```




```{r}
df  <- read.csv("Food Survey - Data.csv")
df
```





###Test 1 - One sample test for mean.


We are using Food Survey as data set and this survey is different kind of people of athlete, non-athlete and gender. We have a total number of observations 167 observation.




```{r}
summary(df$Money)
```



###Below figure is a quantile-quantile plots for a visual inspection of the data normality using Q-Q plots. our Q-Q plot. We can see that the plot is normally distributed for how much money people spend on food.


```{r}
qqnorm(df$Money, pch = 1, frame = FALSE)
qqline(df$Money, col = "steelblue", lwd = 2)

```


```{r}
shapiro.test(df$Money)
```






```{r}
ggboxplot(df$Money, 
          ylab = "Money", xlab = FALSE,
          ggtheme = theme_minimal())
```




###We assume people spend $10 on food so the mu = 10.
H0: Pepole are spending $10 on food.
H1: pepole are not spending $10 on food.

In the result below displays that:



The t is the t-test statistic value (t = 5.6346),
df is the degrees of freedom df = 140,
p-value is the significance level of the t-test p-value < 9.3e-08.
The conf.int is the confidence interval of the mean at 95% (conf.int = [14.82009, 20.03097]);
A sample estimate is the mean value of the sample (mean = 17.42553).



Results
- The p-value is less than 0.005 so at the 5% significance level we reject the null
hypothesis.
- The p-value of the test is less than 9.3e-08, which is less than the significance level alpha
= 0.05. We can conclude that the mean of how much people spend Money in food is significantly different from
10 with a p-value = 9.3e-08.


```{r}
t_test <- t.test(df$Money, mu = 10)
t_test
```




###Below figure is a quantile-quantile plots for a visual inspection of the data normality using Q-Q plots and box . our Q-Q plot. We can see that the plot is normally distributed for NDHMeals.

```{r}
qqnorm(df$NDHMeals, pch = 1, frame = FALSE)
qqline(df$NDHMeals, col = "#0099f8", lwd = 2)

```






```{r}
ggboxplot(df$NDHMeals, 
          ylab = "NDHMeals", xlab = FALSE,
          ggtheme = theme_minimal())
```



###We assume people who had NDHMeals 2 on food so the mu is set to be 2.

H0: Pepole are having 2 NDHMeals
H1: pepole are not having 2 NDHMeals


The t = 4.8558,
df is the degrees of freedom df = 166,
p-value is the significance level of the t-test p-value < 2.752e-06.
With a confidence interval of the mean at 95% conf.int = 2.653808, 3.549785;
The mean = 3.101796.


Here since the since we can find the p value in t.test function so we made another One Sample t-test so that we could extract the p value from there we can view the output below.
The result below shows that the p-value we got is less than 0.005 so at the 5% significance level and it is also greater than the level of significance 0.1. Thus, we concluded that null hypothesis should be rejected at α = 0.05 level of significance.




```{r}
result <- t.test(df$NDHMeals, mu = 2)
result
p_value <- result$p.value
p_value
```





